---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# RMetropolisHastings

<!-- badges: start -->
[![R-CMD-check](https://github.com/rtiwari6-stats/RMetropolisHastings/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/rtiwari6-stats/RMetropolisHastings/actions/workflows/R-CMD-check.yaml)
[![test-coverage](https://github.com/rtiwari6-stats/RMetropolisHastings/actions/workflows/test-coverage.yaml/badge.svg)](https://github.com/rtiwari6-stats/RMetropolisHastings/actions/workflows/test-coverage.yaml)
<!-- badges: end -->

# Description

Metropolis-Hastings (M-H) algorithm is a markov chain based approach that provides a way to generate samples from a distribution from which direct sampling is difficult. It does this by simulating samples from a different distribution from which direct sampling is easier and accepting those samples with a probability. We will build a R package that implements the M-H algorithm.   Our package offers support for univariate and multivariate sampling using the M-H algorithm.


## Installation

You can install the development version of RMetropolisHastings from [GitHub](https://github.com/rtiwari6-stats/RMetropolisHastings) with:

```{r}
if (!require("devtools")){
  install.packages("devtools") 
}
if (!require("RMetropolisHastings")){
  devtools::install_github("rtiwari6-stats/RMetropolisHastings")
}

```


## Examples

This is an example which shows you how to generate 1000 univariate samples:

```{r example}
library(RMetropolisHastings)
targetDensity = function(x){
    return(ifelse(x<0,0,exp(-x)))
}
start = 1.25
y1 = runivariatemh(targetDensity, sigma=1, initial = start, plot = TRUE)
#print a few samples
y1[1:5]
```


This is an example which shows you how to generate 1000 univariate samples using rcpp:

```{r}
library(RMetropolisHastings)
#does not take a userdefined targetDensity
start = -1.25
y1 = cppunivariatemh(targetdensity = "Exponential", sigma=1, initial = start, plot = TRUE)
#print a few samples
y1[1:5]
```

